---
title: "regression_competition"
output: html_document
date: "2024-11-11"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Map of Madrid with points and color indicating prizes

```{r}
library(readxl)
data_train <- read_excel("Data/data_train.xlsx")
```

```{r}
#install.packages(c("ggplot2", "sf", "dplyr"))

# Load libraries
library(ggplot2)
library(sf)
library(dplyr)
library(GGally)

```
# Exploratory Analysis


| **Category** | **Variable**                | **Description**                                                                                                     |
|--------------|-----------------------------|---------------------------------------------------------------------------------------------------------------------|
| Continuous   | precio.house.m2             | Property sale price in euros per square meter                                                                      |
|              | antig                       | Age of the property (years)                                                                                        |
|              | Ruidos_ext                  | External noise (%)                                                                                                 |
|              | Mal_olor                    | Pollution or bad odors (%)                                                                                         |
|              | Poca_limp                   | Lack of street cleanliness (%)                                                                                     |
|              | Malas_comunic               | Poor communications (%)                                                                                            |
|              | Pocas_zonas                 | Few green spaces (%)                                                                                               |
|              | Delincuencia                | Crime (%)                                                                                                          |
|              | CO                          | Level of CO in the air at the property's coordinates (standardized values)                                         |
|              | NO2                         | Level of NO2 in the air at the property's coordinates (standardized values)                                        |
|              | Nox                         | Level of Nox in the air at the property's coordinates (standardized values)                                        |
|              | O3                          | Level of O3 in the air at the property's coordinates (standardized values)                                         |
|              | SO2                         | Level of SO2 in the air at the property's coordinates (standardized values)                                        |
|              | PM10                        | Level of CO in the air at the property's coordinates (standardized values)                                         |
|              | Pobl.0_14_div_Poblac.Total  | Percentage of children between 0 and 14 years in the district                                                     |
|              | PoblJubilada_div_Poblac.Total | Percentage of retired population in the district                                                                 |
|              | Inmigrantes.porc            | Percentage of immigrant population in the district                                                                |
|              | sup.const                   | Built area of the property                                                                                        |
|              | sup.util                    | Usable area of the property                                                                                       |
| Categorical  | barrio/cod_barrio           | Code or name of the neighborhood in the city of Madrid                                                            |
|              | distrito/cod_distrito       | Code or name of the district in the city of Madrid                                                                |
|              | longitud/latitud            | Geographical coordinates of the property (longitude and latitude)                                                 |
|              | dorm                        | Number of bedrooms                                                                                                |
|              | banos                       | Number of bathrooms                                                                                               |
|              | tipo.casa                   | Type of property                                                                                                  |
|              | estado                      | Condition of the property                                                                                         |
| Binary       | inter.exter                 | Interior or exterior design of the property                                                                       |
|              | ascensor                    | Elevator availability in the building                                                                             |
|              | comercial                   | Indicates if the property is located in a commercial area                                                         |
|              | casco.historico             | Indicates if the property is in Madrid's historic center                                                          |
|              | M.30                        | Indicates if the property is within the M-30                                                                      |
| Identifier   | ref.hip.zona                | Mortgage reference of the area                                                                                    |
|              | train_indices               | ID of the observation                                                                                              |

```{r}
colnames(data_train)
```


```{r}
# turn identifiers into strings to prevent errors
data_train[, c("train_indices", "cod_barrio", "cod_distrito", "ref.hip.zona ", "casco.historico", "M.30","comercial")] <- lapply(data_train[, c("train_indices", "cod_barrio", "cod_distrito", "ref.hip.zona", "casco.historico", "M.30","comercial")], as.character)
```


```{r}
unique_counts <- data.frame(
  Column = names(data_train),
  Unique_Values = sapply(data_train, function(x) length(unique(x)))
)
unique_counts
```

```{r}
summary(data_train)
```

## Frequency Analysis

```{r, fig.width=8.27, fig.height=10, out.width='100%'}

plot_hist = function(data_train, type = "both", n, m) {

  # Configure subplots
  par(mfrow = c(n, m),mar = c(10, 4, 4, 1))
  
  # Loop through columns
  for (col in colnames(data_train)) {
    # plot numeric columns
    if (is.numeric(data_train[[col]]) && (type == "numeric" || type == "both")) {
      
      # Create histograms
      hist(data_train[[col]], 
           main = paste(col), 
           xlab = "",
           col = "dodgerblue4")
      
      # plot non-numeric columns
    } else if (!is.numeric(data_train[[col]]) && (type == "categoric" 
                                            || type == "both")) {
      # Adjust x-axis label arrangement for specific cols
      barplot(table(data_train[[col]]), 
              main = paste(col), 
              xlab = "", 
              col = "dodgerblue",
              las = 2)
    }
  }
  
  # Reset the plotting layout
  par(mfrow = c(1, 1))
}

```

```{r, fig.width=8.27, fig.height=10, out.width='100%'}
# Numeric hist
plot_hist(data = data_train ,type="numeric", n=3, m=2)
```

```{r, fig.width=8.27, fig.height=10, out.width='100%'}
# Numeric hist
plot_hist(data = data_train ,type="categoric", n=3, m=2)
```
TODO: Conditional histograms for categoricals (barrios)
```{r warning=FALSE,fig.width=12, fig.height=6, out.width='100%'}
# Compute correlation matrix
# Convert Categorical Variables to Numeric Using Label Encoding
data_encoded <- data_train %>%
  mutate(across(where(is.character), ~ as.numeric(factor(.))))

# Compute the correlation matrix
correlations <- cor(data_encoded)

# Check if 'precio.house.m2' exists
if ("precio.house.m2" %in% colnames(correlations)) {
  # Extract the row corresponding to 'precio.house.m2'
  price_per_sqm_corr <- correlations["precio.house.m2", ]
  
  # Convert to data frame for plotting
  price_per_sqm_corr_df <- data.frame(
    Variable = names(price_per_sqm_corr),
    Correlation = price_per_sqm_corr
  )
  
  # Plot the correlations
  library(ggplot2)
  ggplot(price_per_sqm_corr_df, aes(x = reorder(Variable, Correlation), y = Correlation)) +
    geom_bar(stat = "identity", fill = "steelblue") +
    coord_flip() + # Flip coordinates for easier readability
    theme_minimal() +
    labs(
      title = "Correlation with price_per_sqm",
      x = "Variable",
      y = "Correlation"
    )
} else {
  stop("precio.house.m2 not found in correlation matrix.")
}

```

```{r warning=FALSE,fig.width=30, fig.height=20, out.width='100%'}
library(corrplot)
# Plot
corrplot(correlations, method = "circle", addCoef.col = "black", type = "upper", 
         order = "hclust", tl.col = "black", tl.srt = 45)
```



```{r warning=FALSE,fig.width=30, fig.height=20, out.width='100%'}
# Install and load necessary packages
library(reshape2)
library(ggplot2)
library(GGally) # This is required for ggpairs

# Compute correlation matrix
numeric_column_names <- names(data_train)[sapply(data_train, is.numeric)]
corr_mat <- round(cor(data_train[, numeric_column_names]), 2)
melted_corr_mat <- melt(corr_mat)

# Plot correlation heatmap
ggplot(data = melted_corr_mat, aes(x = Var2, y = Var1, fill = value)) + 
  geom_tile() +
  geom_text(aes(label = value), color = "black", size = 4)+
  scale_fill_gradient(low = "white", high = "red") +
  theme_minimal()

library(ggcorrplot)
p <- ggcorrplot(corr_mat, title = "Matriz de correlaciones",ggtheme = ggplot2::theme_minimal(),
               hc.order = TRUE) + 
  theme(axis.text.x = element_text(size = 7)) + 
  theme(axis.text.y = element_text(size = 7))
ggsave("Matriz de correlaciones.jpg", plot = p, width = 10, height = 10, units = "in", dpi = 300)


# Pairplot using ggpairs
ggpairs(data_train[, numeric_column_names])
```
### Multivariate

```{r warning=FALSE,fig.width=30, fig.height=20, out.width='100%'}
average_price_per_barrio <- data_train %>%
  group_by(barrio) %>%
  summarise(average_price = mean(`precio.house.m2`, na.rm = TRUE))

ggplot(average_price_per_barrio, aes(x = reorder(barrio, average_price), y = average_price)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  coord_flip() +  # Flip coordinates for better readability if there are many barrios
  labs(title = "Average Price per 2m by Barrio",
       x = "Barrio",
       y = "Average Price per 2m") +
  theme_minimal()
```

```{r warning=FALSE,fig.width=20, fig.height=10, out.width='100%'}
average_price_per_district <- data_train %>%
  group_by(distrito) %>%
  summarise(average_price = mean(`precio.house.m2`, na.rm = TRUE))

ggplot(average_price_per_district, aes(x = reorder(distrito, average_price), y = average_price)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  coord_flip() +  # Flip coordinates for better readability if there are many barrios
  labs(title = "Average Price per 2m by District",
       x = "District",
       y = "Average Price per 2m") +
  theme_minimal()
```



```{r}
# creacion de una nueva variable
data_tr_new <- data_train
data_tr_new$sup.const<-NULL
library(geosphere)

centro <- c(-3.6946, 40.4190)  # Longitude, Latitude
data_tr_new$radio <- distHaversine(
  cbind(data_tr_new$longitud, data_tr_new$latitud),
  centro
)
data_tr_new$precio.house.m2 <- log(data_tr_new$precio.house.m2)
numeric_column_names <- names(data_tr_new)[sapply(data_tr_new, is.numeric)]
corr_mat <- round(cor(data_tr_new[, numeric_column_names]), 2)
ggcorrplot(corr_mat, title = "Matriz de correlaciones",ggtheme = ggplot2::theme_minimal(),
               hc.order = TRUE) + 
  theme(axis.text.x = element_text(size = 7)) + 
  theme(axis.text.y = element_text(size = 7))
```
```{r,  fig.width=30, fig.height=20, out.width='100%'}
points_sf <- st_as_sf(data_tr_new, coords = c("longitud", "latitud"), crs = 4326)
points_sf <- points_sf[order(points_sf$precio.house.m2), ]
# Download the map of Madrid (use a simple shapefile, for example)
madrid_map <- st_read("Data/madrid.geojson")
centro <- st_as_sf(data.frame(longitud = -3.6946, latitud = 40.4190), 
                   coords = c("longitud", "latitud"), 
                   crs = 4326)

# Plot the map with the points colored by the 'value' variable
mapa <- ggplot() +
  geom_sf(data = madrid_map, fill = "lightgray", color = "white") +  # Plot map
  geom_sf(data = points_sf, aes(color = precio.house.m2), size = 5) +  # Plot points
  geom_sf(data = centro, color = "yellow", size = 10) + # Plot center
  scale_color_gradient(low = "#00C5CD", high = "red") +
  labs(title = "Spatial situation and price",
       x = "Longitude", y = "Latitude") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate x-axis labels

plot(mapa)
```



```{r,  fig.width=30, fig.height=20, out.width='100%'}
points_sf <- st_as_sf(data_train, coords = c("longitud", "latitud"), crs = 4326)
points_sf <- points_sf[order(points_sf$precio.house.m2), ]
# Download the map of Madrid (use a simple shapefile, for example)
madrid_map <- st_read("Data/madrid.geojson")
centro <- st_as_sf(data.frame(longitud = -3.6946, latitud = 40.4190), 
                   coords = c("longitud", "latitud"), 
                   crs = 4326)

# Plot the map with the points colored by the 'value' variable
mapa <- ggplot() +
  geom_sf(data = madrid_map, fill = "lightgray", color = "white") +  # Plot map
  geom_sf(data = points_sf, aes(color = precio.house.m2), size = 5) +  # Plot points
  geom_sf(data = centro, color = "yellow", size = 10) + # Plot center
  scale_color_gradient(low = "#00C5CD", high = "red") +
  labs(title = "Spatial situation and price",
       x = "Longitude", y = "Latitude") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate x-axis labels

plot(mapa)

### Done with teh help of ChatGPT
```
# Forecasting pipeline

## Preprocess


```{r}

regfit <- regsubsets(precio.house.m2 ~ ., data = data_train,really.big=T)

# Summary of the results
summary(regfit)
```

## Fit & Predict

## Score




