---
title: "regression_competition"
output: html_document
date: "2024-11-11"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Map of Madrid with points and color indicating prizes

```{r}
library(readxl)
data_train <- read_excel("Data/data_train.xlsx")
```

```{r}
#install.packages(c("ggplot2", "sf", "dplyr"))

# Load libraries
library(ggplot2)
library(sf)
library(dplyr)
library(GGally)
library(knitr)
library(kableExtra)
library(geosphere)
library(gridExtra)



```
# Exploratory Analysis


| **Category** | **Variable**                | **Description**                                                                                                     |
|--------------|-----------------------------|---------------------------------------------------------------------------------------------------------------------|
| Continuous   | precio.house.m2             | Property sale price in euros per square meter                                                                      |
|              | antig                       | Age of the property (years)                                                                                        |
|              | Ruidos_ext                  | External noise (%)                                                                                                 |
|              | Mal_olor                    | Pollution or bad odors (%)                                                                                         |
|              | Poca_limp                   | Lack of street cleanliness (%)                                                                                     |
|              | Malas_comunic               | Poor communications (%)                                                                                            |
|              | Pocas_zonas                 | Few green spaces (%)                                                                                               |
|              | Delincuencia                | Crime (%)                                                                                                          |
|              | CO                          | Level of CO in the air at the property's coordinates (standardized values)                                         |
|              | NO2                         | Level of NO2 in the air at the property's coordinates (standardized values)                                        |
|              | Nox                         | Level of Nox in the air at the property's coordinates (standardized values)                                        |
|              | O3                          | Level of O3 in the air at the property's coordinates (standardized values)                                         |
|              | SO2                         | Level of SO2 in the air at the property's coordinates (standardized values)                                        |
|              | PM10                        | Level of CO in the air at the property's coordinates (standardized values)                                         |
|              | Pobl.0_14_div_Poblac.Total  | Percentage of children between 0 and 14 years in the district                                                     |
|              | PoblJubilada_div_Poblac.Total | Percentage of retired population in the district                                                                 |
|              | Inmigrantes.porc            | Percentage of immigrant population in the district                                                                |
|              | sup.const                   | Built area of the property                                                                                        |
|              | sup.util                    | Usable area of the property                                                                                       |
| Categorical  | barrio/cod_barrio           | Code or name of the neighborhood in the city of Madrid                                                            |
|              | distrito/cod_distrito       | Code or name of the district in the city of Madrid                                                                |
|              | longitud/latitud            | Geographical coordinates of the property (longitude and latitude)                                                 |
|              | dorm                        | Number of bedrooms                                                                                                |
|              | banos                       | Number of bathrooms                                                                                               |
|              | tipo.casa                   | Type of property                                                                                                  |
|              | estado                      | Condition of the property                                                                                         |
| Binary       | inter.exter                 | Interior or exterior design of the property                                                                       |
|              | ascensor                    | Elevator availability in the building                                                                             |
|              | comercial                   | Indicates if the property is located in a commercial area                                                         |
|              | casco.historico             | Indicates if the property is in Madrid's historic center                                                          |
|              | M.30                        | Indicates if the property is within the M-30                                                                      |
| Identifier   | ref.hip.zona                | Mortgage reference of the area                                                                                    |
|              | train_indices               | ID of the observation                                                                                              |

```{r}
colnames(data_train)
```


```{r}
# turn identifiers into strings to prevent errors

# data_train[, c("train_indices", "cod_barrio", "cod_distrito", "ref.hip.zona ", "casco.historico", "M.30","comercial")] <- lapply(data_train[, c("train_indices", "cod_barrio", "cod_distrito", "ref.hip.zona", "casco.historico", "M.30","comercial")], as.character)

data_train[, c("train_indices", "cod_barrio", "cod_distrito", "ref.hip.zona")] = lapply(data_train[, c("train_indices", "cod_barrio", "cod_distrito", "ref.hip.zona")], as.character)

cat.var = c("dorm", "banos", "tipo.casa", "inter.exter", "ascensor", "comercial",
            "casco.historico", "M.30")
data_train[, c(cat.var, "barrio", "distrito")] =
  lapply(data_train[, c(cat.var, "barrio", "distrito")], as.factor)

num_id <- sapply(data_train, is.numeric)
num_vars <- names(data_train)[num_id]
```

```{r}
unique_counts <- data.frame(
  Column = names(data_train),
  Unique_Values = sapply(data_train, function(x) length(unique(x)))
)
unique_counts
```

```{r}
summary(data_train)
```

## Frequency Analysis

```{r, fig.width=8.27, fig.height=10, out.width='100%'}

plot_hist = function(data_train, type = "both", n, m) {

  # Configure subplots
  par(mfrow = c(n, m),mar = c(10, 4, 4, 1))
  
  # Loop through columns
  for (col in colnames(data_train)) {
    # plot numeric columns
    if (is.numeric(data_train[[col]]) && (type == "numeric" || type == "both")) {
      
      # Create histograms
      hist(data_train[[col]], 
           main = paste(col), 
           xlab = "",
           col = "dodgerblue4")
      
      # plot non-numeric columns
    } else if (!is.numeric(data_train[[col]]) && (type == "categoric" 
                                            || type == "both")) {
      # Adjust x-axis label arrangement for specific cols
      barplot(table(data_train[[col]]), 
              main = paste(col), 
              xlab = "", 
              col = "dodgerblue",
              las = 2)
    }
  }
  
  # Reset the plotting layout
  par(mfrow = c(1, 1))
}

```

```{r, fig.width=8.27, fig.height=10, out.width='100%'}
# Numeric hist
plot_hist(data = data_train ,type="numeric", n=3, m=2)
```

```{r, fig.width=8.27, fig.height=10, out.width='100%'}
# Numeric hist
plot_hist(data = data_train ,type="categoric", n=3, m=2)
```
TODO: Conditional histograms for categoricals (barrios)
```{r warning=FALSE,fig.width=12, fig.height=6, out.width='100%'}
# Compute correlation matrix
# Convert Categorical Variables to Numeric Using Label Encoding
data_encoded <- data_train %>%
  mutate(across(where(is.character), ~ as.numeric(factor(.))))

# Compute the correlation matrix
correlations <- cor(data_encoded)

# Check if 'precio.house.m2' exists
if ("precio.house.m2" %in% colnames(correlations)) {
  # Extract the row corresponding to 'precio.house.m2'
  price_per_sqm_corr <- correlations["precio.house.m2", ]
  
  # Convert to data frame for plotting
  price_per_sqm_corr_df <- data.frame(
    Variable = names(price_per_sqm_corr),
    Correlation = price_per_sqm_corr
  )
  
  # Plot the correlations
  library(ggplot2)
  ggplot(price_per_sqm_corr_df, aes(x = reorder(Variable, Correlation), y = Correlation)) +
    geom_bar(stat = "identity", fill = "steelblue") +
    coord_flip() + # Flip coordinates for easier readability
    theme_minimal() +
    labs(
      title = "Correlation with price_per_sqm",
      x = "Variable",
      y = "Correlation"
    )
} else {
  stop("precio.house.m2 not found in correlation matrix.")
}

```

```{r warning=FALSE,fig.width=30, fig.height=20, out.width='100%'}
library(corrplot)
# Plot
corrplot(correlations, method = "circle", addCoef.col = "black", type = "upper", 
         order = "hclust", tl.col = "black", tl.srt = 45)
```



```{r warning=FALSE,fig.width=30, fig.height=20, out.width='100%'}
# Install and load necessary packages
library(reshape2)
library(ggplot2)
library(GGally) # This is required for ggpairs

# Compute correlation matrix
numeric_column_names <- names(data_train)[sapply(data_train, is.numeric)]
corr_mat <- round(cor(data_train[, numeric_column_names]), 2)
melted_corr_mat <- melt(corr_mat)

# Plot correlation heatmap
ggplot(data = melted_corr_mat, aes(x = Var2, y = Var1, fill = value)) + 
  geom_tile() +
  geom_text(aes(label = value), color = "black", size = 4)+
  scale_fill_gradient(low = "white", high = "red") +
  theme_minimal()

library(ggcorrplot)
p <- ggcorrplot(corr_mat, title = "Matriz de correlaciones",ggtheme = ggplot2::theme_minimal(),
               hc.order = TRUE) + 
  theme(axis.text.x = element_text(size = 7)) + 
  theme(axis.text.y = element_text(size = 7))
ggsave("Matriz de correlaciones.jpg", plot = p, width = 10, height = 10, units = "in", dpi = 300)


# Pairplot using ggpairs
ggpairs(data_train[, numeric_column_names])

cramersV(data_train$dorm, data_train$banos)

# Initialize a results matrix
cram_results <- matrix(NA, nrow = length(cat.var), ncol = length(cat.var),
                  dimnames = list(cat.var, cat.var))

# Loop through variable pairs
for (i in 1:(length(cat.var))) {
  for (j in (i):length(cat.var)) {
    var1 <- cat.var[i]
    var2 <- cat.var[j]
    
    # Calculate Cramér's V
    cram_results[i, j] <- cramersV(data_train[[var1]], data_train[[var2]])
    cram_results[j, i] <- cram_results[i, j]  # Symmetric
  }
}

# Convert to a heatmap-friendly format
print(cram_results)

# Melt results for heatmap
melted_results <- melt(cram_results, na.rm = TRUE)

# Plot heatmap
ggplot(melted_results, aes(Var1, Var2, fill = value)) +
  geom_tile(color = "white") +  # Add white borders for better contrast
  geom_text(aes(label = sprintf("%.2f", value)), size = 12, color = "black") +  # Larger text for values
  scale_fill_gradient(low = "white", high = "#CD5C5C", limits = c(0, 1)) +  # Adjust color scale
  theme_minimal() +
  theme(
    axis.text.x = element_text(size = 20, angle = 45, hjust = 1),  # Larger x-axis labels
    axis.text.y = element_text(size = 20),  # Larger y-axis labels
    plot.title = element_text(size = 24, face = "bold", hjust = 0.5),  # Larger title
    legend.title = element_text(size = 18),  # Larger legend title
    legend.text = element_text(size = 16)  # Larger legend text
  ) +
  labs(title = "Cramér's V Heatmap", fill = "Cramér's V", x = "", y = "")

```

### Objective vs. Predictors

In this part we will be creating scatterplots of the relation between the continuous predictors and their price.

```{r, warning=FALSE}

num_vars_2 = setdiff(num_vars, c("precio.house.m2"))#, "ref.hip.zona"))

for (i in num_vars_2) {
  
  # Calculate R^2
  fit <- lm(precio.house.m2 ~ data_train[[i]], data = data_train)
  r_squared <- summary(fit)$r.squared
  
  # Create the plot
  p <- ggplot(data = data_train, aes(x = data_train[[i]], y = precio.house.m2)) +
    geom_point(color = "darkseagreen4", alpha = 0.6) +  # Scatter plot
    geom_smooth(method = "lm", se = FALSE, color = "#FFAEB9") +  # Regression line
    theme_bw() +
    labs(
      title = paste0("Variable:", i, " (R² = ", round(r_squared, 3), ")"),
      x = i,
      y = "precio.house.m2"
    )
  plot(p)
}
```

We observe poor correlation between the objective value and the predictor features.

### Categorical Values Selection

Let's consider the relations that exist between the categorical values and the price of the flats

```{r}

# 
# summary(data_train[cat.var])
# 
# # Define a function to calculate and display contingency tables with proportions
# explore_categorical_interactions <- function(data, cat_vars) {
#   for (i in 1:(length(cat_vars) - 1)) {
#     for (j in (i + 1):length(cat_vars)) {
#       var1 <- cat_vars[i]
#       var2 <- cat_vars[j]
#       
#       # Create contingency table (frequencies)
#       freq_table <- table(data[[var1]], data[[var2]])
#       
#       # Create proportion table
#       prop_table <- 100*prop.table(freq_table) %>% round(4)
#       
#       # Combine frequency and proportion into one data frame
#       combined_table <- cbind("Absolute:"= freq_table,
#                               "Relative:" = prop_table)
# 
#       # Display using kableExtra
#       cat("\n### Interaction Between", var1, "and", var2, "\n\n")
#       combined_table %>% print()
#     }
#   }
# }
# 
# 
# explore_categorical_interactions(data_train, cat.var)
```

Now lets consider the interaction with the price

```{r}
# Define a function to create boxplots for price grouped by categorical variables
plot_price_by_categories <- function(data, cat_vars, price_var) {
  for (var in cat_vars) {
    # Create the boxplot
    p <- ggplot(data, aes_string(x = var, y = price_var)) +
      geom_boxplot(fill = "#69b3a2", color = "black", outlier.color = "red") +
      theme_bw() +
      labs(
        title = paste("Boxplot of", price_var, "by", var),
        x = var,
        y = price_var
      ) +
      theme(
        axis.text.x = element_text(angle = 45, hjust = 1, size = 12),
        axis.text.y = element_text(size = 12),
        plot.title = element_text(size = 16, face = "bold", hjust = 0.5)
      )
    
    # Print the plot
    print(p)
  }
}

# Example usage
plot_price_by_categories(data_train, cat.var, "precio.house.m2")
```

```
### Multivariate

```{r warning=FALSE,fig.width=30, fig.height=20, out.width='100%'}
average_price_per_barrio <- data_train %>%
  group_by(barrio) %>%
  summarise(average_price = mean(`precio.house.m2`, na.rm = TRUE))

ggplot(average_price_per_barrio, aes(x = reorder(barrio, average_price), y = average_price)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  coord_flip() +  # Flip coordinates for better readability if there are many barrios
  labs(title = "Average Price per 2m by Barrio",
       x = "Barrio",
       y = "Average Price per 2m") +
  theme_minimal()
```

```{r warning=FALSE,fig.width=20, fig.height=10, out.width='100%'}
average_price_per_district <- data_train %>%
  group_by(distrito) %>%
  summarise(average_price = mean(`precio.house.m2`, na.rm = TRUE))

ggplot(average_price_per_district, aes(x = reorder(distrito, average_price), y = average_price)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  coord_flip() +  # Flip coordinates for better readability if there are many barrios
  labs(title = "Average Price per 2m by District",
       x = "District",
       y = "Average Price per 2m") +
  theme_minimal()
```



```{r}
# creacion de una nueva variable
data_tr_new <- data_train
data_tr_new$sup.const<-NULL
library(geosphere)

centro <- c(-3.6946, 40.4190)  # Longitude, Latitude
data_tr_new$radio <- distHaversine(
  cbind(data_tr_new$longitud, data_tr_new$latitud),
  centro
)
# logarithmic objective
data_tr_new$log.precio.house.m2 <- log(data_tr_new$precio.house.m2)
numeric_column_names <- names(data_tr_new)[sapply(data_tr_new, is.numeric)]
corr_mat <- round(cor(data_tr_new[, numeric_column_names]), 2)
ggcorrplot(corr_mat, title = "Matriz de correlaciones",ggtheme = ggplot2::theme_minimal(),
               hc.order = TRUE) + 
  theme(axis.text.x = element_text(size = 7)) + 
  theme(axis.text.y = element_text(size = 7))
```
```{r,  fig.width=30, fig.height=20, out.width='100%'}
points_sf <- st_as_sf(data_tr_new, coords = c("longitud", "latitud"), crs = 4326)
points_sf <- points_sf[order(points_sf$precio.house.m2), ]
# Download the map of Madrid (use a simple shapefile, for example)
madrid_map <- st_read("Data/madrid.geojson")
# Central point (Puerta del Sol)
center <- c(-3.7038, 40.4168)

# Calculate distances and add a new column
data_train$radius <- distHaversine(
  matrix(c(data_train$longitud, data_train$latitud), ncol = 2),
  matrix(rep(center, nrow(data_train)), ncol = 2, byrow = TRUE)
) / 1000  # Convert meters to kilometers


```

```{r warning=FALSE,fig.width=30, fig.height=20, out.width='100%'}
# Function to create a ggplot map with a specified column as the color variable
create_district_map <- function(points_sf, madrid_map, color_column) {
  column_data <- points_sf[[color_column]]


  ggplot() +
    # Base map of Madrid
    geom_sf(data = madrid_map, fill = "lightgray", color = "white") +
    
    # Points representing districts or other categories
    geom_sf(data = points_sf, aes_string(color = color_column), size = 3, alpha = 0.8) +
    
    # Title and axis labels
    labs(
      title = "Geographical Distribution by Districts",
      x = "Longitude",
      y = "Latitude",
      color = color_column  # Dynamically set legend title
    ) +
    
    # Improved theme
    theme_minimal() +
    theme(
      plot.title = element_text(size = 2*30, face = "bold", hjust = 0.5),
      plot.subtitle = element_text(size = 22, hjust = 0.5),
      axis.title.x = element_text(size = 2*20),
      axis.title.y = element_text(size = 2*20),
      axis.text.x = element_text(size = 2*16, angle = 45, hjust = 1),
      axis.text.y = element_text(size = 2*16),
      legend.title = element_text(size = 3*20),
      legend.text = element_text(size = 3*16),
      legend.key.size = unit(4, "cm"),
      panel.grid.major = element_line(color = "gray80", linetype = "dashed"),
      panel.grid.minor = element_blank()
    )+
    
      if (is.numeric(column_data) && length(unique(column_data)) > 10) {
            scale_color_viridis_c(option = "turbo")
      } else {
            scale_color_viridis_d(option = "turbo")  # Discrete color scale
      }


}

# Loop through all columns in points_sf
columns = names(points_sf)
columns <- setdiff(columns, c("train_indices", "geometry"))
points_sf$ref.hip.zona <- as.numeric(points_sf$ref.hip.zona)


# for (col_name in columns) {
#   # Create the plot for the current column
#   plot <- create_district_map(points_sf, madrid_map, col_name)
#   
#   # Print the plot
#   print(plot)
# }


```




# Specific Questions Corner

## Grouping Barrio/Distrito

```{r warning=FALSE,fig.width=30, fig.height=20, out.width='100%'}




points_sf <- points_sf %>%
    mutate(
      distrito = case_when(
        # South Districts
        distrito %in% c("arganzuela" ,"centro", "chamartin", "chamberi", "moncloa", "retiro", "salamanca") ~ "Center",
        
        # Central Districts
        distrito %in% c("carabanchel", "latina") ~ "South West",
        
        # North Districts
        distrito %in% c("moratalaz", "puente_vallecas", "usera", "vallecas", "vicalvaro", "villaverde") ~ "South East",
        
        # West Districts
        distrito %in% c("barajas","ciudad_lineal", "fuencarral", "hortaleza", "san_blas","tetuan") ~ "North East",
        
        # Default to original values if no match
        TRUE ~ as.character(distrito)
      )
    )
points_sf$distrito <- factor(points_sf$distrito, levels = unique(points_sf$distrito))


plot <- create_district_map(points_sf, madrid_map, "distrito")
print(plot)

```


```{r warning=FALSE,fig.width=20, fig.height=10, out.width='100%'}
average_price_per_district <- data_train %>%
  group_by(distrito) %>%
  summarise(average_price = mean(`precio.house.m2`, na.rm = TRUE))

ggplot(average_price_per_district, aes(x = reorder(distrito, average_price), y = average_price)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  coord_flip() +  # Flip coordinates for better readability if there are many barrios
  labs(title = "Average Price per 2m by District",
       x = "District",
       y = "Average Price per 2m") +
  theme_minimal()


average_price_per_barrio <- data_train %>%
  group_by(barrio) %>%
  summarise(average_price = mean(`precio.house.m2`, na.rm = TRUE))

ggplot(average_price_per_barrio, aes(x = reorder(barrio, average_price), y = average_price)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  coord_flip() +  # Flip coordinates for better readability if there are many barrios
  labs(title = "Average Price per 2m by Barrio",
       x = "Barrio",
       y = "Average Price per 2m") +
  theme_minimal()

```



