---
title: "regression_competition"
output: html_document
date: "2024-11-11"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Map of Madrid with points and color indicating prizes

```{r}
library(readxl)
data_train <- read_excel("Data/data_train.xlsx")
```

```{r}
#install.packages(c("ggplot2", "sf", "dplyr"))

# Load libraries
library(ggplot2)
library(sf)
library(dplyr)
library(GGally)

```

# Exploratory Analysis

| **Category** | **Variable**                  | **Description**                                                             |
|-------------------|-------------------|-----------------------------------|
| Continuous   | precio.house.m2               | Property sale price in euros per square meter                               |
|              | antig                         | Age of the property (years)                                                 |
|              | Ruidos_ext                    | External noise (%)                                                          |
|              | Mal_olor                      | Pollution or bad odors (%)                                                  |
|              | Poca_limp                     | Lack of street cleanliness (%)                                              |
|              | Malas_comunic                 | Poor communications (%)                                                     |
|              | Pocas_zonas                   | Few green spaces (%)                                                        |
|              | Delincuencia                  | Crime rate (%)                                                              |
|              | CO                            | Level of CO in the air at the property's coordinates (standardized values)  |
|              | NO2                           | Level of NO2 in the air at the property's coordinates (standardized values) |
|              | Nox                           | Level of Nox in the air at the property's coordinates (standardized values) |
|              | O3                            | Level of O3 in the air at the property's coordinates (standardized values)  |
|              | SO2                           | Level of SO2 in the air at the property's coordinates (standardized values) |
|              | PM10                          | Level of CO in the air at the property's coordinates (standardized values)  |
|              | Pobl.0_14_div_Poblac.Total    | Percentage of children between 0 and 14 years in the district               |
|              | PoblJubilada_div_Poblac.Total | Percentage of retired population in the district                            |
|              | Inmigrantes.porc              | Percentage of immigrant population in the district                          |
|              | sup.const                     | Built area of the property                                                  |
|              | sup.util                      | Usable area of the property                                                 |
| Categorical  | barrio/cod_barrio             | Code or name of the neighborhood in the city of Madrid                      |
|              | distrito/cod_distrito         | Code or name of the district in the city of Madrid                          |
|              | longitud/latitud              | Geographical coordinates of the property (longitude and latitude)           |
|              | dorm                          | Number of bedrooms                                                          |
|              | banos                         | Number of bathrooms                                                         |
|              | tipo.casa                     | Type of property                                                            |
|              | estado                        | Condition of the property                                                   |
| Binary       | inter.exter                   | Interior or exterior design of the property                                 |
|              | ascensor                      | Elevator availability in the building                                       |
|              | comercial                     | Indicates if the property is located in a commercial area                   |
|              | casco.historico               | Indicates if the property is in Madrid's historic center                    |
|              | M.30                          | Indicates if the property is within the M-30                                |
| Identifier   | ref.hip.zona                  | Mortgage reference of the area                                              |
|              | train_indices                 | ID of the observation                                                       |

```{r}
colnames(data_train)
```

### Preprocessing

```{r}
# turn identifiers into strings to prevent errors

# data_train[, c("train_indices", "cod_barrio", "cod_distrito", "ref.hip.zona ", "casco.historico", "M.30","comercial")] <- lapply(data_train[, c("train_indices", "cod_barrio", "cod_distrito", "ref.hip.zona", "casco.historico", "M.30","comercial")], as.character)

data_train[, c("train_indices", "cod_barrio", "cod_distrito", "ref.hip.zona ")] = lapply(data_train[, c("train_indices", "cod_barrio", "cod_distrito", "ref.hip.zona")], as.character)

cat.var = c("dorm", "banos", "tipo.casa", "inter.exter", "ascensor", "comercial",
            "casco.historico", "M.30")
data_train[, cat.var] = lapply(data_train[, cat.var], as.factor)

num_id <- sapply(data_train, is.numeric)
num_vars <- names(data_train)[num_id]
```

```{r}
unique_counts <- data.frame(
  Column = names(data_train),
  Unique_Values = sapply(data_train, function(x) length(unique(x)))
)
unique_counts
```

```{r}
summary(data_train)
```

## Frequency Analysis

```{r, fig.width=8.27, fig.height=10, out.width='100%'}

plot_hist = function(data_train, type = "both", n, m) {

  # Configure subplots
  par(mfrow = c(n, m),mar = c(10, 4, 4, 1))
  
  # Loop through columns
  for (col in colnames(data_train)) {
    # plot numeric columns
    if (is.numeric(data_train[[col]]) && (type == "numeric" || type == "both")) {
      
      # Create histograms
      hist(data_train[[col]], 
           main = paste(col), 
           xlab = "",
           col = "dodgerblue4")
      
      # plot non-numeric columns
    } else if (!is.numeric(data_train[[col]]) && (type == "categoric" 
                                            || type == "both")) {
      # Adjust x-axis label arrangement for specific cols
      barplot(table(data_train[[col]]), 
              main = paste(col), 
              xlab = "", 
              col = "dodgerblue",
              las = 2)
    }
  }
  
  # Reset the plotting layout
  par(mfrow = c(1, 1))
}
```

```{r, fig.width=8.27, fig.height=10, out.width='100%'}
# Numeric hist
plot_hist(data = data_train ,type="numeric", n=3, m=2)
```

```{r, fig.width=8.27, fig.height=10, out.width='100%'}
# Numeric hist
plot_hist(data = data_train ,type="categoric", n=3, m=2)
```

TODO: Conditional histograms for categoricals (barrios)

```{r warning=FALSE,fig.width=12, fig.height=6, out.width='100%'}
# Compute correlation matrix
# Convert Categorical Variables to Numeric Using Label Encoding
data_encoded <- data_train %>%
  mutate(across(where(is.character), ~ as.numeric(factor(.))))

# Compute the correlation matrix
correlations <- cor(data_encoded)

# Check if 'precio.house.m2' exists
if ("precio.house.m2" %in% colnames(correlations)) {
  # Extract the row corresponding to 'precio.house.m2'
  price_per_sqm_corr <- correlations["precio.house.m2", ]
  
  # Convert to data frame for plotting
  price_per_sqm_corr_df <- data.frame(
    Variable = names(price_per_sqm_corr),
    Correlation = price_per_sqm_corr
  )
  
  # Plot the correlations
  library(ggplot2)
  ggplot(price_per_sqm_corr_df, aes(x = reorder(Variable, Correlation), y = Correlation)) +
    geom_bar(stat = "identity", fill = "steelblue") +
    coord_flip() + # Flip coordinates for easier readability
    theme_minimal() +
    labs(
      title = "Correlation with price_per_sqm",
      x = "Variable",
      y = "Correlation"
    )
} else {
  stop("precio.house.m2 not found in correlation matrix.")
}

```

```{r warning=FALSE,fig.width=30, fig.height=20, out.width='100%'}
library(corrplot)
# Plot
corrplot(correlations, method = "circle", addCoef.col = "black", type = "upper", 
         order = "hclust", tl.col = "black", tl.srt = 45)

```

### Correlation metrics

Lets check the correlation that exist between the variables, for the continuous variable we will compute the (Pearson's) correlation matrix and for the discrete ones we will consider the Cramér's V.

$$ V = \sqrt{\frac{\chi^2}{n \cdot \min(r-1, c-1)}} \quad \text{where} \quad \chi^2 = \sum_{i,j} \frac{\left(n_{ij} - \frac{n_{i.} n_{.j}}{n}\right)^2}{\frac{n_{i.} n_{.j}}{n}}$$

Cramér's V measures the association between two categorical variables, ranging from 0 (no association) to 1 (perfect association).

It is derived from the chi-squared statistic ($\chi^2$), which measures how far away the observed frequencies in a contingency table differ from the expected frequencies under the assumption of no association, it is adjusted for the total number of observations ($n$) and the degrees of freedom ($\text{min}(r-1, c-1)$) to account for table size and structure. The square root scales is due to the nature of the chi-squared statistic measure is squared.

The values of interpretation variate greatly on the context of the data but as a rule of thumb (CHATGPT) we have:

-   $0\leq V \leq 0.1$: Weak association.

-   $0.1 \le V \le 0.3$: Moderate association.

-   $0.3 \le V \le 1$: Strong association.

```{r warning=FALSE,fig.width=30, fig.height=20, out.width='100%'}
# Install and load necessary packages
library(reshape2)
library(ggplot2)
library(GGally) # This is required for ggpairs
library(lsr)

# Compute correlation matrix
numeric_column_names <- names(data_train)[sapply(data_train, is.numeric)]
corr_mat <- round(cor(data_train[, numeric_column_names]), 2)
melted_corr_mat <- melt(corr_mat)

# Plot correlation heatmap with middle white color for 0
ggplot(data = melted_corr_mat, aes(x = Var2, y = Var1, fill = value)) + 
  geom_tile() +
  geom_text(aes(label = value), color = "black", size = 4) +
  scale_fill_gradient2(
    low = "#4A708B",    # Blue for low correlations
    mid = "white",      # White for 0 correlations
    high = "#CD5C5C",   # Red for high correlations
    midpoint = 0,       # Center the color scale at 0
    limits = c(-1, 1),  # Optional: specify limits for the color scale
    name = "Correlation"  # Optional: add a legend title
  ) +
  theme_minimal() +
  labs(title = "Correlation Heatmap",
       x = "Variables",
       y = "Variables")

# Pairplot using ggpairs
ggpairs(data_train[, numeric_column_names])

cramersV(data_train$dorm, data_train$banos)

# Initialize a results matrix
cram_results <- matrix(NA, nrow = length(cat.var), ncol = length(cat.var),
                  dimnames = list(cat.var, cat.var))

# Loop through variable pairs
for (i in 1:(length(cat.var))) {
  for (j in (i):length(cat.var)) {
    var1 <- cat.var[i]
    var2 <- cat.var[j]
    
    # Calculate Cramér's V
    cram_results[i, j] <- cramersV(data_train[[var1]], data_train[[var2]])
    cram_results[j, i] <- results[i, j]  # Symmetric
  }
}

# Convert to a heatmap-friendly format
print(cram_results)

```

### Objective vs. Predictors

In this part we will be creating scatterplots of the relation between the continuous predictors and their price.

```{r, warning=FALSE}

num_vars_2 = setdiff(num_vars, c("precio.house.m2", "ref.hip.zona"))

for (i in num_vars_2) {
  
  # Calculate R^2
  fit <- lm(precio.house.m2 ~ data_train[[i]], data = data_train)
  r_squared <- summary(fit)$r.squared
  
  # Create the plot
  p <- ggplot(data = data_train, aes(x = data_train[[i]], y = precio.house.m2)) +
    geom_point(color = "darkseagreen4", alpha = 0.6) +  # Scatter plot
    geom_smooth(method = "lm", se = FALSE, color = "#FFAEB9") +  # Regression line
    theme_bw() +
    labs(
      title = paste0("Variable:", i, " (R² = ", round(r_squared, 3), ")"),
      x = i,
      y = "precio.house.m2"
    )
  plot(p)
}
```

We observe poor correlation between the objective value and the predictor features.

### Multivariate

```{r warning=FALSE,fig.width=30, fig.height=20, out.width='100%'}
average_price_per_barrio <- data_train %>%
  group_by(barrio) %>%
  summarise(average_price = mean(`precio.house.m2`, na.rm = TRUE))

ggplot(average_price_per_barrio, aes(x = reorder(barrio, average_price), y = average_price)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  coord_flip() +  # Flip coordinates for better readability if there are many barrios
  labs(title = "Average Price per 2m by Barrio",
       x = "Barrio",
       y = "Average Price per 2m") +
  theme_minimal()
```

```{r warning=FALSE,fig.width=20, fig.height=10, out.width='100%'}
average_price_per_district <- data_train %>%
  group_by(distrito) %>%
  summarise(average_price = mean(`precio.house.m2`, na.rm = TRUE))

ggplot(average_price_per_district, aes(x = reorder(distrito, average_price), y = average_price)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  coord_flip() +  # Flip coordinates for better readability if there are many barrios
  labs(title = "Average Price per 2m by District",
       x = "District",
       y = "Average Price per 2m") +
  theme_minimal()
```

```{r warning=FALSE,fig.width=30, fig.height=20, out.width='100%'}
points_sf <- st_as_sf(data_train, coords = c("longitud", "latitud"), crs = 4326)
points_sf <- points_sf[order(points_sf$precio.house.m2), ]
# Download the map of Madrid (use a simple shapefile, for example)
madrid_map <- st_read("Data/madrid.geojson")

# Plot the map with the points colored by the 'value' variable
ggplot() +
  geom_sf(data = madrid_map, fill = "lightgray", color = "white") +  # Plot map
  geom_sf(data = points_sf, aes(color = precio.house.m2), size = 4) +  # Plot points
  scale_color_gradient(low = "#00C5CD", high = "red") +
  labs(title = "Spatial situation and price",
       x = "Longitude", y = "Latitude") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate x-axis labels


### Done with teh help of ChatGPT
```

# Forecasting pipeline

## Preprocess

```{r}

# regfit <- regsubsets(precio.house.m2 ~ ., data = data_train,really.big=T)
# 
# # Summary of the results
# summary(regfit)
```

## Fit & Predict

## Score
